---
title: "class_regex"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
#install.packages("stringr")
library("stringr")
library("dplyr")
library("ggplot2")
```


## Basic string manipulaiton
```{r}
test_str <- c("STAT545 is great!", "Wooooo!")
nchar(test_str)

#Extract the 1st through 7th elements
substr(test_str, 1, 7)

#Replace the 1st through 7th elemebt with SCIENCE
substr(test_str, 1, 7) <- "SCIENCE"
test_str
```

## paste - concatenate or combine strings
```{r}
#default is a single space
paste("abc", "efg")

#can define a separator
paste("abc", "efg", sep = "")
#or eqivalently
paste0("abc", "efg")

#to combine multi value vectors, also need collapse
paste(c("abc", "efg"), c("hij", "klm"), sep = "", collapse = "")
```

## strsplit - split a string into a list of substrings

```{r}
x <- c("abc,cbe", "cb,gb,aaa")
#split based on commas
strsplit(x, split = ",")

#that returns a list, so perhaps we only want every second element of each
strsplit(x, split = ",") %>% lapply(function(x) x[2])
```

## Regex in R

R has some regex that makes it more readable. Can also use the \\w etc equivalents. Examples:

* \w = [[:alnum:]]
* \d = [[:digit:]]
* \s = [[:space:]]

You often need to double escape, eg \\\\w

## Load some tweets to experiment with

```{r}
news_tweets <- read.delim("https://www.dropbox.com/s/cbgcpkizun51wbk/news_tweets.txt?dl=1", header = TRUE, stringsAsFactors = FALSE, sep = "\t", quote = "", allowEscapes = TRUE)
glimpse(news_tweets)
```

* user_name = twitter user name (shortened)
* created = date tweeted
* retweets = number of retweets to date
* text = the raw text of the tweet

## grep - find a pattern in a character vector
```{r}
#find tweets that contain hashtags
# two identical ways of writing this regex:
hastag_pattern <- "#\\w+"
hastag_pattern <- "#[[:alnum:]]+"

#gives the row number of the tweet that contains a hashtag
grep(hastag_pattern, news_tweets$text)

#Setting value = TRUE returns the actual value of the vector at those indexes:
grep(hastag_pattern, news_tweets$text, value = TRUE) %>% head(n = 5)

#cal also return the ones that do not contain has tag with invert=TRUE
grep(hastag_pattern, news_tweets$text, value = TRUE, invert = TRUE) %>% head(n = 5)
```





