---
title: "class_dplyr"
author: "Lauren Fratamico"
date: "September 29, 2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Install and load dplyr and load gapminder:

```{r}
suppressPackageStartupMessages(library(dplyr)) #suppresses the output
library(gapminder)
```

Let's look at some gapminder data

```{r}
str(gapminder)
head(gapminder)
```


Turn the data frame into a tbl_df. This is not necessary for dplyr, but it allows it some nice functions.
```{r}
gtbl <- tbl_df(gapminder)
gtbl
glimpse(gtbl)
```


Can store subsets of data
```{r}
(snippet <- subset(gapminder, country == "Canada"))
```

## Filtering

filter is useful for subsetting data row-wise
```{r}
filter(gtbl, lifeExp < 29)
filter(gtbl, country == "Rwanda")
filter(gtbl, country %in% c("Rwanda", "Afghanistan"))
```

In base R, we could filter/subset like so:
```{r}
gapminder[gapminder$lifeExp < 29, ]
subset(gapminder, country == "Rwanda")
```


## Pipe object

The pipe object pipes the left as the first argument to the right. More useful when chaining function calls - more readable.
```{r}
gapminder %>% head #same as head(gapminder)
gapminder %>% head(3)
```


Select with the pipe to get just the head
```{r}
select(gtbl, year, lifeExp)
gtbl %>%
  select(year, lifeExp) %>%
  head(4)
```

More practive with pipe
```{r}
gtbl %>%
  filter(country == "Cambodia") %>%
  select(year, lifeExp)

# typical R command would look like:
gapminder[gapminder$country == "Cambodia", c("year", "lifeExp")]
# or like:
subset(gapminder, country == "Cambodia", select = c(year, lifeExp))
```




